---
title: "Simulation studies for mis-classification in non-probability surveys"
author: "Maciej Beręsewicz"
format: 
  html:
    self-contained: true
    toc: true
---

## Basic setup

Read the packages for the simulation studies

```{r setup, message=FALSE, warning=FALSE}
library(simex) ## only confusion matrix is provided
library(augSIMEX) ## more advanced methods
library(misclassGLM)
library(nonprobsvy)
library(data.table)
```

## Simulation studies

### Simulation study #1

This is the first simulation study for the project. In this simulation study I check the performance of MC-SIMEX method for mis-classification where at least one auxiliary variable is measured with error. Then I verify the case when the target variable is also measured with error. 

#### Simulation with mis-classified auxiliary variables 

##### Description

Here is the description for the simulation study

+ $N=100000$ -- finite population size
+ $X_1$ -- categorical variable with 3 levels: `c("S", "M", "L")` with probabilities `c(0.7, 0.2, 0.1)`. This variable may refer to the size of the business entity.
+ $X_2 \sim \text{Bernoulli}(0.7)$ -- categorical variable referring to business sector (say 1=private, 0=public).
+ $X_3 \sim \text{Bernoulli}(0.5)$ -- categorical variable referring to some characteristic that is equally distributed in the population.

I create the target variable $Y$ from Bernoulli distribution with probabilities generated from the logistic distribution where linear part is given by 

$$
-1 + 0.5 \times I(X_1 = \text{L}) + 0.5 \times I(X_2 = \text{Public}) + -0.3 \times X3,
$$
where $I(\cdot)$ is an indicator variable. So the target variable is related with all variables considered in the simulation study.

I generate two samples:

+ non-probability sample $S_A$ with inclusion generated from Bernoulli distribution with probabilities generated from the logistic distribution with linear part given by

$$
-4 + 3 \times I(X_1 = \text{L})
$$
and where the expected sample size is about 4,500.

+ probability sample $S_B$ with 1,000 sample size selected using SRS.

Next, for each simulation run I introduce classification errors for $X_1$ and $X_2$ given the following confusion matrices which elements are $\pi_{ij} = P(X^* == k | X == k)$ where $k$ are levels

$$
\Gamma_{X_1} = \begin{bmatrix}
0.8 & 0.1 & 0.05 \\
0.15 & 0.7 &  0.15\\
0.05 &  0.2 &  0.8
\end{bmatrix}
$$
and

$$
\Gamma_{X_2} = 
\begin{bmatrix}
0.80 & 0.15\\
0.20 & 0.85
\end{bmatrix}
$$
where rows and columns in $\Gamma_{X_1}$ and $\Gamma_{X_2}$ refer to levels i.e. "S", "M", "L" and  Private and Public respectively. So it means that in each simulation run instead of $X_1$ and $X_2$ we observe $X_1^*$ and $X_2^*$ in non-probability sample.



For this simulation study I am interested in $\mu_Y$ which is population mean (share) of $Y$. In the simulation study I compare the following estimators:

+ naive estimator $\mu_{Y}^{\text{Naive}}$ which is a sample mean based on non-probability sample,
+ two estimators when no mis-classification is observed in non-probability sample:
  + inverse probability weighted (IPW) estimator with calibration constraints $\mu_{Y}^{IPW}$ on $X_1$, $X_2$ and $X_3$ where calibration is performed with totals estimated from the probability sample,
  + mass imputation (MI) estimator based on logistic regression $\mu_Y^{MI}$ using all auxiliary variables $X_1$, $X_2$ and $X_3$
+ two estimators when mis-classification is observed for non-probability sample *only*:
  + IPW with mis-classification in $X_1$ and $X_2$ which is denoted as $\mu_{Y}^{IPW*}$
  + MI with mis-classification in $X_1$ and $X_2$ which is denoted as $\mu_{Y}^{MI*}$
+ mass imputation estimator where mis-classification in $X_1$ and $X_2$ is corrected using MC-SIMEX method assuming that the confusion matrix is known. This estimator is denoted as $\mu_{Y}^{MC-MI}$. 

##### Codes to run the simulation

I set up the confusion matrices

```{r}
## rows - X*, columns -- X (true)
Gamma_X1 <- matrix(c(0.8,  0.1, 0.05,
                     0.15, 0.7, 0.15,
                     0.05, 0.2, 0.8), nrow = 3, byrow = T)
  
rownames(Gamma_X1) <- colnames(Gamma_X1) <- c("S", "M", "L")
  
Gamma_X2 <- matrix(c(0.8, 0.15,
                     0.2, 0.85), byrow = T, ncol = 2)

rownames(Gamma_X2) <- colnames(Gamma_X2) <- c("Pub", "Priv")

```

In the code below I generate population data

```{r population-data}
set.seed(12345)
N <- 100000  ## population size
n_b <- 1000 ## probability sample
X1 <- sample(c("S", "M", "L"), N, prob = c(0.6, 0.3, 0.1), replace = T) ## X1 variable
X1 <- factor(X1, c("S", "M", "L"))
X2 <- rbinom(N, 1, prob = 0.7) ## X2 variable
X2 <- factor(X2, 0:1, c("Pub", "Priv")) 
X3 <- rbinom(N, 1, prob = 0.5) ## X3 variable 
Y <- rbinom(N, 1, plogis(1 + 5*(X1 == "L") - 3*(X2 == "Pub"))) ## target variable
pi_A <- plogis(-2 + 2*(X1 == "L") + (X2 == "Priv")) ## selection to non-probability sample ~ 20%
pop_data <- data.table(X1,X2,X3,Y,pi_A) ## population data frame
head(pop_data)
```

The known population mean of $Y$ is `r round(mean(Y)*100,2)`%.


Code to run the simulation

```{r simulation-codes}
n_reps <- 100
results <- list()

for (b in 1:n_reps) {
  set.seed(b)
  print(b)
  ## probability sample
  sample_b <- pop_data[sample(1:N, n_b), ]
  sample_b$sample_w <- N/n_b
  sample_b_svy <- svydesign(ids=~1, weights = ~ sample_w, data=sample_b)
  
  ## nonprobability sample
  sample_a <- pop_data[rbinom(N, 1, pop_data$pi_A) == 1, ]
  
  # without error -----------------------------------------------------------
  est_ipw <- tryCatch(nonprob(selection = ~ X1 + X2 + X3,
                     data = sample_a,
                     svydesign = sample_b_svy,
                     target = ~ Y), error = function(e) "error")
  if (class(est_ipw)[1] == "character") next
  
  est_mi <- nonprob(outcome = Y ~ X1 + X2 + X3,
                    data = sample_a,
                    svydesign = sample_b_svy, 
                    family_outcome = "binomial")
  
  ## with errors
  sample_a_with_error <- copy(sample_a)
  sample_a_with_error[X2 == "Pub",  X2_star := factor(sample(c("Pub", "Priv"), size=.N, prob=Gamma_X2[, 1], replace = T), levels = c("Pub", "Priv"))]
  sample_a_with_error[X2 == "Priv", X2_star := factor(sample(c("Pub", "Priv"), size=.N, prob=Gamma_X2[, 2], replace = T), levels = c("Pub", "Priv"))]
  sample_a_with_error[X1 == "S",    X1_star := factor(sample(c("S", "M", "L"), size=.N, prob=Gamma_X1[,1], replace = T), levels = c("S", "M", "L"))]
  sample_a_with_error[X1 == "M",    X1_star := factor(sample(c("S", "M", "L"), size=.N, prob=Gamma_X1[,2], replace = T), levels = c("S", "M", "L"))]
  sample_a_with_error[X1 == "L",    X1_star := factor(sample(c("S", "M", "L"), size=.N, prob=Gamma_X1[,3], replace = T), levels = c("S", "M", "L"))]
  sample_a_with_error[, X1:=factor(X1_star, levels = c("S", "M", "L"))]
  sample_a_with_error[, X2:=factor(X2_star, levels = c("Pub", "Priv"))]
  
  est_ipw_err <- tryCatch(nonprob(selection = ~ X1 + X2 + X3,
                         data = sample_a_with_error,
                         svydesign = sample_b_svy,
                         target = ~ Y), error = function(e) "error")
  
  if (class(est_ipw_err)[1] == "character") next
  
  est_mi_err <- nonprob(outcome = Y ~ X1 + X2 + X3,
                        data = sample_a_with_error,
                        svydesign = sample_b_svy, 
                        family_outcome = "binomial")
  
  ## naive model

  est_mi_err_naive <- glm(Y ~ X1 + X2 + X3, family = binomial, x = TRUE, y = TRUE,
                          data = sample_a_with_error)
  
  ## model parameters corrected by MC-SIMEX approach
  res_mi_q <- simex::mcsimex(model = est_mi_err_naive,
                           mc.matrix = list(X1 = Gamma_X1, X2 = Gamma_X2),
                           SIMEXvariable = c("X1", "X2"),
                           fitting.method	= "quadratic",
                            asymptotic = FALSE,
                            jackknife.estimation	= FALSE,
                            B = 100,
                            lambda = c(0.01, 0.05, 0.10, 0.5, 1))
  
  res_mi_l <- refit(res_mi_q, fitting.method	= "linear", asymptotic = FALSE,
                            jackknife.estimation	= FALSE)
  
  sample_b_svy <- update(sample_b_svy, 
         y_mi_q = predict(res_mi_q, sample_b_svy$variables, type = "response"),
         y_mi_l = predict(res_mi_l, sample_b_svy$variables, type = "response"))
  
  est <- svymean(~y_mi_q+y_mi_l, design = sample_b_svy)
  est
  
  results[[b]] <- data.frame(b, 
                             mu_naive = mean(sample_a$Y),
                             mu_ipw = est_ipw$output$mean[1], 
                             mu_mi = est_mi$output$mean[1],
                             mu_ipw_star = est_ipw_err$output$mean[1], 
                             mu_mi_star = est_mi_err$output$mean[1],
                             mu_mc_mi_q = est[1],
                             mu_mc_mi_l = est[2])
  
}

results_df <- rbindlist(results)
```

Boxplots with the simulation results

```{r}
boxplot((results_df[,-1]-mean(pop_data$Y))/mean(pop_data$Y)*100)
abline(a=0,b=0,h=0,col='red')
```
Bias, SD and RMSE
```{r}
results <- rbind(
  colMeans(results_df[,-1]) - mean(Y),
  apply(results_df[,-1], 2, sd),
  sqrt((colMeans(results_df[,-1]) - mean(Y))^2 + apply(results_df[,-1], 2, var)) 
)
rownames(results) <- c("bias", "SD", "RMSE")
as.data.frame(results)
```




```{r}
class_erros <- misclass(data.frame(X1 = sample_a_with_error$X1,
                                   X2 = sample_a_with_error$X2), 
                        list(X1=Gamma_X1, X2=Gamma_X2), k = 1)

table(sample_a_with_error$X1, class_erros$X1) |> prop.table(margin = 1)
```

```{r}
est_mi_err <- nonprob(outcome = Y ~ X1 + X2 + X3,
                      data = sample_a_with_error,
                      svydesign = sample_b_svy, 
                      family_outcome = "binomial")

mm <- misclassGLM()
```

